### 2013-08-29 pystruct/paper.tex
FIX put chapter commands in separate files, so that the text is not on a new page...
Fixes in the cutting plane algorithms, nicer looks

@@

\section{Learning Algorithms for Linear Max-Margin Structured Prediction}
Maximum Margin learning is now one of the most popular methods to learn
structural models in computervision and in other fields.
There are several reasons for the popularity of linear maximum margin approaches:
\begin{description}
    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
        minimize a user-specified loss.
    \item[Feasibility] If the loss decomposes over the factor graph that
        specifies $g$, then learning is feasible as soon as the maximization
        over $y$ in \Eqref{main_equation} can be caried out.
    \item[Generalization] The maximum margin principle yields generalization
        bounds using the effective complexity~\citep{taskar2003max}.
    \item[Strong Convexity] The resulting optimization problem is strongly
        convex, leading to efficient optimization and unique solutions.
\end{description}

For learning, a dataset $(x^{(1)}, y^{(1)}),\dotsc,(x^{(k)}, y^{(k)})$, together with a loss
\begin{equation}
    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
\end{equation}
The parameters $\theta$ are learned by minimizing the loss based soft-margin
objective
\begin{equation}\eqlabel{learning_equation}
    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x^{(i)}, y^{(i)}, \theta)
\end{equation}
with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
on $\Delta$ empirical risk:
\begin{equation}\eqlabel{loss_augmentation}
    \ell(x^{(i)}, y^{(i)}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y^{(i)}, y) + \theta^T \phi(x^{(i)}, y) - \theta^T \phi(x^{(i)}, y^{(i)}]_+
\end{equation}

@@

\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
into a quadratic objective with a combinatorial number of constraints:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
is maintained, using the cutting plane method. The algorithm starts with an empty working set,
and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
again, with the new set of constraints.
The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
which guarantees a suboptimality of at most $\epsilon$.

The complete algorithm is described in Algorithm~\ref{alg_n_slack}.

%\begin{algorithm*}[t]
    %\caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
    %%FIXME

    %\begin{algorithmic}[1]
        %\require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
        %\ensure parameters $\theta$, slack $\xi$
        %\state $\w \leftarrow \emptyset$
        %\repeat
        %\state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
        %\begin{equation*}
            %\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
            %\left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
                %- \xi
            %\end{equation*}
            %\for {i=1, \dots, n}
            %\state
            %$\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
            %\endfor
            %\state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
            %\state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
            %\until $\xi' - \xi < \epsilon$ \label{convergence_check}
        %\end{algorithmic}
    %\end{algorithm*}

@@

\begin{algorithm*}[t]
    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
    \begin{algorithmic}[1]
        \require training samples $\{ (x^{(1)}, y^{(1)}), \dots, (x^{(n)}, y^{(n)})\}$, regularization parameter $c$, stopping tolerance $\epsilon$.
        \ensure parameters $\theta$, slack $\xi$
        \state $\w \leftarrow \emptyset$
        \repeat
            \state $(\theta, \xi) \leftarrow \displaystyle \argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + c \xi$\label{restricted}\\
            \begin{equation*}
                \text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \in \w:
                \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^i, \hat{y}^{(i)})] \right \rangle \geq \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)})
                - \xi
            \end{equation*}
            \for {i=1, \dots, n}
                \state
                $\hat{y}^{(i)} \leftarrow i(x^{(i)}, y^{(i)}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{y}} \sum_{i=1}^n \delta(y^{(i)}, \hat{y}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^{(i)}, y^{(i)}) - \psi(x^{(i)}, \hat{y})] \right \rangle$ \label{get_cutting_plane}
            \endfor
            \state $\w \leftarrow \w \cup \{ (\hat{y}^{(1)}, \dots, \hat{y}^{(n)}) \} $
            \state $ \displaystyle \xi' \leftarrow  \sum_{i=1}^n \delta(y^{(i)}, \hat{y}^{(i)}) - \left \langle \theta, \sum_{i=1}^n [\psi(x^i, y^i) - \psi(x^{(i)}, \hat{y}^{(i)})] \right \rangle$
        \until $\xi' - \xi < \epsilon$ \label{convergence_check}
    \end{algorithmic}
\end{algorithm*}

The one-slack cutting plane method~\citep{joachims2009cutting} solves the
following reformulation of Equation~\Eqref{learning_equation}:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
Informally, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points, and
then applying the n-slack algorithm.
The complete algorithm is described in Algorithm~\ref{alg_one_slack}.

>>>>

\chapter{Concepts and Implementation of Conditional Random Field-based Semantic Segmentation}\label{ch:structured_pystruct}

@@

\section{Learning Algorithms for Linear Max-Margin Structured Prediction}

Maximum Margin learning is now one of the most popular methods to learn
structural models in computervision and in other fields.
There are several reasons for the popularity of linear maximum margin approaches:
\begin{description}
    \item[Loss-sensitivity] In contrast to probabilistic approaches, maximum margin learning approaches can directly
        minimize a user-specified loss.
    \item[Feasibility] If the loss decomposes over the factor graph that
        specifies $g$, then learning is feasible as soon as the maximization
        over $y$ in \Eqref{main_equation} can be caried out.
    \item[Generalization] The maximum margin principle yields generalization
        bounds using the effective complexity~\citep{taskar2003max}.
    \item[Strong Convexity] The resulting optimization problem is strongly
        convex, leading to efficient optimization and unique solutions.
\end{description}

For learning, a dataset $(x\hoch{1}, y\hoch{1}),\dotsc,(x\hoch{k}, y\hoch{k})$ is given, together with a loss
\begin{equation}
    \Delta \colon \mathcal{Y} \times \mathcal{Y} \rightarrow \mathbb{R}.
\end{equation}
The parameters $\theta$ are learned by minimizing the loss based soft-margin
objective
\begin{equation}\eqlabel{learning_equation}
    \min_\theta \frac{1}{2} ||\theta||^2 + C \sum_i  \ell(x\hoch{i}, y\hoch{i}, \theta)
\end{equation}
with regularization parameter $C$. Here $r_i$ is an hinge-loss like upper bound
on $\Delta$ empirical risk:
\begin{equation}\eqlabel{loss_augmentation}
    \ell(x\hoch{i}, y\hoch{i}, \theta)= [\max_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y) - \theta^T \phi(x\hoch{i}, y\hoch{i}]_+
\end{equation}

@@

\subsection{The N-Slack Cutting Plane Method}\seclabel{n_slack}
\begin{algorithm*}[t]
    \caption{N-Slack Cutting Plane Training of Structural SVMs \label{alg_n_slack}}
    \begin{doublespacing}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{1}, y\hoch{1}), \dots, (x\hoch{k}, y\hoch{k})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $(\xi_1, \dotsc, \xi_k)$
        \State $\W_i \leftarrow \emptyset, \xi_i \leftarrow 0 \text{ for all } i=1,\dots,k$
        \Repeat
            \For {i=1, \dots, k}
                \State
                $\hat{y} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] $
                \If{$\displaystyle \delta(y\hoch{i}, \hat{y}) - \theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})] \geq \xi_i + \epsilon$}
                    \State $\W_i \leftarrow \W_i \cup \{ \hat{y}\} $
                    \State
                    \vspace{-15mm}
                    \begin{align*}
                    (\theta, \xi_1, \dots, \xi_k) \leftarrow \displaystyle &\argmin_{\theta, \xi_1, \dots, \xi_k} \frac{||\theta||}{2}^2 + C \sum_{i=1}^k\xi_i\\
                    &\text{s.t. }\forall i=1,\dots,k \forall \hat{y} \in \W_i:\\
                    &\theta^T [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi_i
                    \end{align*}
                \EndIf
            \EndFor
            \vspace{-10mm}
            \Until no $\W_i$ changes any more.
        \end{algorithmic}
    \end{doublespacing}
    \end{algorithm*}

The n-slack cutting plane method~\citep{tsochantaridis2006large} reformulates \Eqref{learning_equation}
into a quadratic objective with a combinatorial number of constraints:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \sum_{i=1}^k \xi_i\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \mathcal{Y}^k:\\
    &\theta^T \sum_{i=1}^n [\phi(x\hoch{i}, y\hoch{i}) - \phi(x\hoch{i},
        \hat{y}\hoch{i})] \geq \sum_{i=1}^k \Delta(y\hoch{i}, \hat{y}\hoch{i})
            - \xi
\end{align}
As is not feasible to deal with all constraints, only a working set $\W$ of active constraints
is maintained, using the cutting plane method. The algorithm starts with an empty working set,
and in each iteration adds the most violated constraint to $\W$. Then, the quadratic program is solved
again, with the new set of constraints.
The algorithm terminates when no constraint can be found that is more violated then $\epsilon$,
which guarantees a suboptimality of at most $\epsilon$.

The complete procedure is described in Algorithm~\ref{alg_n_slack}.

@@

\begin{algorithm*}[t]
    \caption{1-Slack Cutting Plane Training of Structural SVMs \label{alg_one_slack}}
    \begin{algorithmic}[1]
        \Require training samples $\{ (x\hoch{i}, y\hoch{i}), \dots, (x\hoch{i}, y\hoch{i})\}$, regularization parameter $C$, stopping tolerance $\epsilon$.
        \Ensure parameters $\theta$, slack $\xi$
        \State $\W \leftarrow \emptyset$
        \Repeat
            \State 
            \vspace{-5mm}
            \begin{align*}
            (\theta, \xi) \leftarrow \displaystyle &\argmin_{\theta, \xi} \frac{||\theta||}{2}^2 + C \xi\\
            &\text{s.t. }\forall \hat{\mathbf{y}}=(\hat{y}\hoch{1}, \dots, \hat{y}\hoch{k}) \in \W:\\
            &\theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] \geq \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \xi
            \end{align*}
            \For {i=1, \dots, k}
                \State
                $\hat{y}\hoch{i} \leftarrow I(x\hoch{i}, y\hoch{i}, \theta) := \displaystyle \argmax_{\hat{y}\in\mathcal{Y}} \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y})]$ \label{get_cutting_plane}
            \EndFor
            \State $\W \leftarrow \W \cup \{ (\hat{y}\hoch{i}, \dots, \hat{y}\hoch{i}) \} $
            \State $ \displaystyle \xi' \leftarrow  \sum_{i=1}^k \delta(y\hoch{i}, \hat{y}\hoch{i}) - \theta^T \sum_{i=1}^k [\psi(x\hoch{i}, y\hoch{i}) - \psi(x\hoch{i}, \hat{y}\hoch{i})] $
        \Until $\xi' - \xi < \epsilon$ \label{convergence_check}
    \end{algorithmic}
\end{algorithm*}

The one-slack cutting plane method~\citep{joachims2009cutting} solves the
following reformulation of Equation~\Eqref{learning_equation}:
\begin{align}\label{eq:oneslack}
    \min_{\theta, \xi}\ &\frac{1}{2} ||\theta||^2 + C \xi\\
    \text{s.t. }&\forall \hat{\mathbf{y}}=(\hat{y}^1, \dots, \hat{y}^n) \in \mathcal{Y}^n:\\
        &\theta^T \sum_{i=1}^n [\phi(x^i, y^i) - \phi(x^i,
            \hat{y}^i)] \geq \sum_{i=1}^n \Delta(y^i, \hat{y}^i)
            - \xi
\end{align}
Informally, the one-slack formulation corresponds to joining all training
samples into a single training example $(\mathbf{x}, \mathbf{y})$ that has no
interactions between variables corresponding to different data points, and
then applying the n-slack algorithm.
A detailed description can be found in Algorithm~\ref{alg_one_slack}.
